Live Functional Programming with Typed Holes

### 2018-02-26 calculus.tex
fig-dynamics-aux
fig-dynamics-contexts
metatheory definitions, wip

\input{fig-substitution}
\input{fig-dynamics-aux}
\input{fig-dynamics-steps}
\input{fig-dynamics-contexts}

@@

@@

\clearpage
\subsection{Metatheory \rkc{TODO}}

(these theorems should be taken as fairly provisional)

\begin{theorem}[Expandability \cy{??}] ~
  \begin{enumerate}
    \item If $\hsyn{\hGamma}{\hexp}{\htau}$ then $\expandSyn{\hGamma}{\hexp}{\htau}{\dexp}{\Delta}$ for some $\dexp$ and $\Delta$.
    \item If $\hana{\hGamma}{\hexp}{\htau}$ then $\expandAna{\hGamma}{\hexp}{\htau}{\dexp}{\htau'}{\Delta}$ for some $\dexp$ and $\htau'$ and $\Delta$.
  \end{enumerate}
\end{theorem}

\begin{theorem}[Correspondence \cy{??}] ~
  \begin{enumerate}
    \item If $\expandSyn{\hGamma}{\hexp}{\htau}{\dexp}{\Delta}$ then $\hsyn{\hGamma}{\hexp}{\htau}$.
    \item If $\expandAna{\hGamma}{\hexp}{\htau}{\dexp}{\htau'}{\Delta}$ then $\hana{\hGamma}{\hexp}{\htau}$.
  \end{enumerate}
\end{theorem}

@@

\begin{theorem}[Progress]
If $\hasType{\Delta}{\emptyset}{\dexp}{\htau}$ then either $\stepsToD{\Delta}{\dexp}{\dexp'}$ or 
$\isValue{\dexp}$ or $\isErr{\Delta}{\dexp}$ or $\isIndet{\dexp}$. 
% for some $\dexp'$.
\end{theorem}

\begin{theorem}[Preservation]
  If $\hasType{\Delta}{\emptyset}{\dexp}{\htau}$ and
  $\stepsToD{\Delta}{\dexp}{\dexp'}$ then
  $\hasType{\Delta}{\emptyset}{\dexp'}{\htau}$. \cy{or maybe up to
  consistency?}
\end{theorem}

@@

>>>>

\input{fig-dynamics-aux}
%% \input{fig-dynamics-steps}
\input{fig-dynamics-contexts}

@@

\begin{lemma}[Focus Formation] ~
  If $\selectEvalCtx{\dexp}{\evalctx}{\dexp'}$
  then $\isevalctx{\evalctx}$.
\end{lemma}

\begin{lemma}[Matched Ground Type Invariant] ~
  If $\groundmatch{\htau}{\htau'}$
  then $\isGround{\htau'}$
  and $\tconsistent{\htau}{\htau'}$
  and $\htau\neq\htau'$.
\end{lemma}

@@

\clearpage
%% \subsection{Metatheory}
\subsection{Metatheory: Type Soundness}

\begin{theorem}[Expandability] ~
  \begin{enumerate}
    \item
      If $\hsyn{\hGamma}{\hexp}{\htau}$
      then $\expandSyn{\hGamma}{\hexp}{\htau}{\dexp}{\Delta}$
      for some $\dexp$ and $\Delta$.
    \item
      If $\hana{\hGamma}{\hexp}{\htau}$
      then $\expandAna{\hGamma}{\hexp}{\htau}{\dexp}{\htau'}{\Delta}$
      for some $\dexp$ and $\htau'$ and $\Delta$.
  \end{enumerate}
\end{theorem}

\begin{theorem}[Correspondence] ~
  \begin{enumerate}
    \item
      If $\expandSyn{\hGamma}{\hexp}{\htau}{\dexp}{\Delta}$
      then $\hsyn{\hGamma}{\hexp}{\htau}$.
    \item
      If $\expandAna{\hGamma}{\hexp}{\htau}{\dexp}{\htau'}{\Delta}$
      then $\hana{\hGamma}{\hexp}{\htau}$.
  \end{enumerate}
\end{theorem}

@@

\begin{theorem}[Preservation]
  If $\hasType{\Delta}{\emptyset}{\dexp}{\htau}$ and
  $\stepsToD{\Delta}{\dexp}{\dexp'}$ then
  $\hasType{\Delta}{\emptyset}{\dexp'}{\htau}$.
\end{theorem}

\begin{theorem}[Progress]
  If $\hasType{\Delta}{\emptyset}{\dexp}{\htau}$ then either
  (i) $\stepsToD{}{\dexp}{\dexp'}$ or
  (ii) $\isIndet{\dexp}$ or
  (iii) $\isBoxedValue{\dexp}$.
\end{theorem}

\begin{theorem}[Canonical Value Forms]
  If $\hasType{\hDelta}{\emptyset}{\dexp}{\htau}$ and $\isValue{\dexp}$
  then $\htau\neq\tehole$ and
  \begin{enumerate}[label=(\alph*)]
    \item If $\htau=b$ then $\dexp=c$.
    \item If $\htau=\tarr{\htau_1}{\htau_2}$
          then $\dexp=\halam{x}{\htau_1}{\dexp'}$
          where $\hasType{\hDelta}{x : \htau_1}{\dexp'}{\htau_2}$.
  \end{enumerate}
\end{theorem}

\begin{theorem}[Canonical Boxed Forms]
  If $\hasType{\hDelta}{\emptyset}{\dexp}{\htau}$ and $\isBoxedValue{\dexp}$
  then
  \begin{enumerate}[label=(\alph*)]
    \item If $\htau=b$ then $\dexp=c$.
    \item If $\htau=\tarr{\htau_1}{\htau_2}$ then either
      \begin{enumerate}
        \item[i.]
          $\dexp=\halam{x}{\htau_1}{\dexp'}$
          where $\hasType{\hDelta}{x : \htau_1}{\dexp'}{\htau_2}$, or
        \item[ii.]
          $\dexp=\dcasttwo{\dexp'}{\tarr{\htau_1'}{\htau_2'}}{\tarr{\htau_1}{\htau_2}}$
          where $\tarr{\htau_1'}{\htau_2'}\neq\tarr{\htau_1}{\htau_2}$
          and $\hasType{\hDelta}{\emptyset}{\dexp'}{\tarr{\htau_1'}{\htau_2'}}$.
      \end{enumerate}
    \item If $\htau=\tehole$
          then $\dexp=\dcasttwo{\dexp'}{\htau'}{\tehole}$
          where $\isGround{\htau'}$
          and $\hasType{\hDelta}{\emptyset}{\dexp'}{\htau'}$.
  \end{enumerate}
\end{theorem}

\begin{theorem}[Canonical Indeterminate Forms]
  \rkc{TODO}

  \rkc{TODO}

  \rkc{TODO}

  \rkc{TODO}

  \rkc{TODO}

  \rkc{TODO}

  \rkc{TODO}

  \rkc{TODO}
\end{theorem}

\input{fig-complete}

\autoref{fig:complete} defines complete types and expressions...

@@

\subsection{Metatheory: Correctness of Edit-and-Resume}

\input{fig-substitution}
